Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces

### 2011-10-25 TransitOrientation.tex
Final resubmission to the journal

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ of the transiting \suldiox~(Figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface location, resulting in the 2-dimensional angle and depth-profile histograms shown in Figure \ref{fig:so2-transit-angles}. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/transit-so2-angles/so2-angles-transit.png}
		\caption{\suldiox~molecular orientation distributions during SMD transit simulations into an aqueous slab. Both the neat-water (top row) and saturated (bottom row) data sets were analyzed to determine the angles $\theta$ (left column), and $\phi$ (right column) of the transiting \suldiox. The distributions show angle cosines plotted against the distance of the \suldiox~to the location of the water surface. The data sets were averaged at each distance for the respective 50 simulations, and every distribution at each distance was normalized such that each histogram has a max population of 1 for purposes of comparison.}
		\label{fig:so2-transit-angles}
	\end{center}
\end{figure}

	From its starting position above the water surface, until the \suldiox~moves to within 10\angs~of the water surface location of both systems, the orientation is isotropic in $\theta$ and $\phi$ (not shown). Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$. Near and into the interfacial region the bisector angle $\theta$ becomes more perpendicular to the water surface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above. At the point when the \suldiox~reaches the water surface location (distance $=0$\angs), the bisector is perpendicular to the interface in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the neat and saturated water systems is where the point of the transition from isotropic to preferred orientations is found.

  % talk about theta differences
  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface water location. Below 5\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system. It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation. From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
		\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface. The \suldiox~sulfur points down towards the aqueous phase. (B) In the neat-water system, a transiting \suldiox~molecule located below the water surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
		\label{fig:so2-surface-angles}
	\end{center}
\end{figure}

  % talk about phi differences
	%Indeed, the $\phi$ orientations appear mostly isotropic without any clear trends for preferred orientation in that angle. This appears as a mostly uniform color across the entire plot, with minor variations that are disjointed throughout the distance range.

 With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis. This is the case in both systems, with only a few exceptions. In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend. At the neat-water surface and just below, from 0\angs~to -5\angs~(in the water phase), the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot. This corresponds to \suldiox~inclined up to 60 degrees from the surface normal. The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{fig:so2-surface-angles}B depicts the orientation of the \suldiox~slightly below the water surface location. This is in contrast to \suldiox~molecules above the water surface location, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface location. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions in the interfacial region. %Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.

>>>>

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ of the transiting \suldiox~(Figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The orientation depth-profiles were collected for the 50 simulations of both systems for various distances from the water surface location, resulting in the 2-dimensional angle and depth-profile histograms shown in Figure \ref{fig:so2-transit-angles}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles/theta-phi-transit.png}
		\caption{Molecular orientation distributions of an \suldiox~at different interfacial depths of an aqueous slab during SMD transit simulations. Both the neat-water (top) and saturated (bottom) data sets were analyzed to determine the angles $\theta$ (horizontal axes) and $\phi$ (vertical axes) of the transiting \suldiox. The distributions shown are those for the single transiting \suldiox~molecule averaged over the 50 SMD simulations.}
		\label{fig:so2-transit-angles}
	\end{center}
\end{figure}

	From its starting position above the water surface, until the \suldiox~moves to within 6\angs~of the water surface location of both systems, the orientation is isotropic in $\theta$ and $\phi$. Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\theta$ or $\phi$. Near and into the interfacial region the bisector angle $\theta$ becomes more perpendicular to the water surface ($\theta \approx 0$\textdegree) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above. At the point when the \suldiox~reaches the water surface location (0\angs), the bisector is perpendicular to the interface in both the neat-water and saturated systems. %Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. 
	In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the neat and saturated water systems is where the point of the transition from isotropic to preferred orientations is found.

  % theta differences
  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 4\angs~above the surface water location. Below 4\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system. It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation. From onset of orientation above the surface until 10\angs~below (not shown), the \suldiox~holds a preferred orientation.

%\begin{figure}[h!]
	%\begin{center}
		%\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
		%\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface. The \suldiox~sulfur points down towards the aqueous phase. (B) In the neat-water system, a transiting \suldiox~molecule located below the water surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
		%\label{fig:so2-surface-angles}
	%\end{center}
%\end{figure}

  % phi differences
	%Indeed, the $\phi$ orientations appear mostly isotropic without any clear trends for preferred orientation in that angle. This appears as a mostly uniform color across the entire plot, with minor variations that are disjointed throughout the distance range.

 With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis. This is the case in both systems, with only a few exceptions. In both systems the $\phi$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend. At the neat-water surface and just above (from 0-2\angs~above the water phase), the $\theta$ profile broadens to $\theta = 90$\textdegree, near $\phi = 90$\textdegree~appearing as a shoulder of light coloration in the bottom-left of the 0-2\angs~axes. This indicates that \suldiox~inclined up to 90\textdegree~from the surface normal will have a preferred $\phi$ orientation lying more flat to the water surface. This is in contrast to \suldiox~molecules above the water surface location, oriented more perpendicularly and without a preference for a particular range of values in $\phi$. The behavior is likely due to the interaction between the waters and the S-O bonds leading to a higher solvation than above the water surface location. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions in the interfacial region. %Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
